\section{Conclusion and Future Work}
\label{sec:concl}

We present an method for computing Craig interpolants for first-order formulas over real numbers with a wide range of nonlinear functions.
Our method transform proof traces from δ decision procedures into interpolants consisting of disjunctive linear constraints.
The algorithms are guaranteed to find the interpolants between two formulas $A$ and $B$ whenever $A ∧ B$ is not δ-satisfiable.
Furthermore, we show how the framework apply to systems of ordinary differential equations.
We implemented our interpolation algorithm in the \dReal SMT-solver and apply the method to domains such robotic design, and hybrid system verification.  

In the future, we plan to expand our work to richer proof systems.
The ICP loop produces proof based on interval pruning which results in large, ``squarish'' interpolants.
Using more general proof systems, e.g. cutting planes, we will be able to get smaller, smoother interpolants.
CDCL-style reasoning for richer theories, e.g., LA(ℝ)~\cite{DBLP:conf/cav/McMillanKS09} and polynomial~\cite{DBLP:conf/cade/JovanovicM12}, is a likely basis for such extensions.
Furthermore, we are interested in investigating the link between classical interpolation and Craig interpolation over the reals.
Using methods like spline interpolation and radial basis functions, it maybe possible to build smoother interpolants.
%We also to extend the our rules to compute interpolants mixed proofs with both integer and real variables.
